---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# noisecanceling

<!-- badges: start -->
<!-- badges: end -->

The goal of `noisecanceling` is to run basic psychometric analysis of scales composed of likert scaled items implementing acquiescence correction. It assumes a test composed of likert item types with some semantic pairs. This package have also some functions that facilitates runing psychometric analysis of tests composed of multiple scales.


## Installation

You can install the development version of noisecanceling from [GitHub](https://github.com/) with:

```{r}
# install.packages("devtools")
devtools::install_github("rprimi/noisecanceling")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}

 library(noisecanceling)
 library(magrittr)

 senna_recoded <-data %>% 
    select(unique(item_dic$coditem)) %>% 
    recode_for_acq( 
    item_dic = item_dic
  )

 senna_psicom <- find_psychometrics(senna_recoded, likert = 5, center = 3)
  
 save_item_psicom( senna_psicom , filename = "senna_psicom.xlsx") 
  

```


