---
title: "Correcting questionnaires for acquiescence bias"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Correcting questionnaires for acquiescence bias}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE
)
```

```{r setup}
library(noisecanceling)
```

## Acquiescence as noise

**Acquiescence** is the tendency to agree with questionnaire items regardless
of their content. It adds a person-specific offset to every response and
distorts scale scores, reliabilities and correlations.

`noisecanceling` treats acquiescence the way noise-cancelling headphones treat
ambient sound. The headphones record the noise and play back an inverted copy
(180° out of phase); the noise and its inverted copy cancel, leaving the
signal. In a **balanced scale**:

- the **trait** is the signal;
- **acquiescence** is the noise, and it pushes responses *upward* on both
  positively keyed (PK) and negatively keyed (NK) items;
- reverse-scoring the NK items realigns the trait signal while leaving
  acquiescence split into equal positive and negative parts that cancel.

The package estimates a per-person acquiescence index from the balanced items
and centers responses on it, so the remaining variance is a purified measure
of the trait.

## The item dictionary

Every function is driven by an **item dictionary**: a data frame with one row
per item describing how it should be scored. The bundled `senna_dic` is an
example.

```{r}
data(senna_dic)
head(senna_dic[, c("coditem", "scale", "pole", "seman_pairs", "item_text")])
```

The required columns are:

| Column        | Meaning                                                        |
|---------------|----------------------------------------------------------------|
| `coditem`     | Item code; matches a column name in the response data.         |
| `scale`       | Scale the item is scored on.                                   |
| `pole`        | `1` = positively keyed, `0` = negatively keyed.                |
| `seman_pairs` | Identifier of the PK/NK semantic pair; `NA` if the item is not paired. |
| `item_text`   | Item text (optional, carried through for reporting).           |

Items with a non-missing `seman_pairs` value are the **balanced items** used to
estimate acquiescence.

## Step 1 — Recode for acquiescence

`recode_for_acq()` computes the acquiescence index (the within-person mean of
all paired items) and centers every response on it.

```{r}
data(data_senna)

recoded <- recode_for_acq(data_senna, senna_dic)
str(recoded, max.level = 1)
```

`acq_index` holds the per-person index and within-subject standard deviation:

```{r}
head(recoded$acq_index)
```

`data_acq_recoded` holds the centered responses — each value is the original
response minus that person's acquiescence index.

## Step 2 — Classical psychometrics

`find_psychometrics()` reverse-scores the NK items and runs a classical
psychometric analysis (scale scores, Cronbach's alpha, item-total
correlations) on **both** the original and the acquiescence-corrected data.

```{r}
psicom <- find_psychometrics(recoded, likert = 5, center = 3)
```

Compare the reliabilities side by side:

```{r}
orig <- psicom$alpha_orig_scale_stat[, c("scale", "raw_alpha")]
rec  <- psicom$alpha_rec_scale_stat[, c("scale", "raw_alpha")]
merge(orig, rec, by = "scale", suffixes = c("_original", "_recoded"))
```

The `scores` element contains the scale scores (`_ori` and `_rec`) together
with the acquiescence indices, ready for further analysis.

```{r}
head(psicom$scores)
```

## Scoring without an acquiescence object

`score_tests()` scores scales directly from a data frame and an item
dictionary, which is handy for scoring either raw or already-recoded data.

```{r}
scored <- score_tests(data_senna, senna_dic)
scored$alpha_scale_stat[, c("scale", "raw_alpha", "std.alpha")]
```

## Exporting results

`save_item_psicom()` writes the full `find_psychometrics()` output to a
multi-sheet Excel workbook, and `save_item_psicom_si()` does the same for
`score_tests()` output:

```{r, eval = FALSE}
save_item_psicom(psicom, filename = "senna_psychometrics.xlsx")
```

## Summary

1. Build an **item dictionary** describing scales, poles and semantic pairs.
2. `recode_for_acq()` — estimate acquiescence and center the responses.
3. `find_psychometrics()` or `score_tests()` — analyse the corrected scores.
4. `save_item_psicom()` — export the results.

See the [function reference](../reference/index.html) for the full API.
