---
title: "exemplo_simples"
author: "Ricardo Primi"
date: "9/21/2019"
output: html_document
---


#### Simulating balanced scale of 4 items (2 positively keyed and 2 negatively keyed)  

* Produces a dataset 6 items of 5 point likers scale.  
* Create all$5^6=15625$ possible response patterns  



```{r echo=TRUE}
  
  library(crossdes)
  library(tidyverse)

  r_ptrn <- permutations(5, 4, repeats.allowed=TRUE)
  colnames(r_ptrn) <- paste("i", 1:4, sep="")
  
 

```
 
* Create mean and standard deviation within subjects and recode responses using two methods: (a) centering $c_{ij} = x_{ij}-acq_j$ and (b) $z_{ij} = (x_{ij}-acq_j)/sd_j$  


```{r echo=TRUE}
 
  acq_j <- apply(r_ptrn, 1, mean)
  sd_j <- apply(r_ptrn, 1, sd)
  
  r_ptrn_c <- apply(r_ptrn, 2, function(x) { x - acq_j} )
  r_ptrn_z <- apply(r_ptrn, 2, function(x) { (x - acq_j)/sd_j} )
  

```

* Calculate scores. Note we reverse three last items as if they were negatively worded items.

```{r echo=TRUE}

  scr_j  <- r_ptrn %>% 
    as.data.frame %>%
    mutate_at(3:4,  funs(6-.)) %>%
    apply(1, mean, na.rm=TRUE)
  
  scr.rec_j  <- r_ptrn_c %>% 
    as.data.frame %>%
    mutate_at(3:4,  funs(-1*.)) %>%
    apply(1, mean, na.rm=TRUE)
  
  scr.z.rec_j  <- r_ptrn_z %>% 
    as.data.frame %>%
    mutate_at(3:4,  funs(-1*.)) %>%
    apply(1, mean, na.rm=TRUE)

```


* Create a tibble with scores acquiescence index  

```{r echo=TRUE}

 df <- tibble(
   scr_j,
   scr.rec_j,
   scr.z.rec_j,
   acq_j,
   sd_j
  )

```


#### Visualizing relationship among scores and acquiescence parameters

* How recoded scores (standardized and centered) are related to original scores
* How sd_j infuences transformation
* What is the relaioship between acq_j (mean) and standard deviation sd_j 



```{r, echo=TRUE}
  
  library(artyfarty)
  library(psych)
  library(knitr)
  library(ggplot2)
  library(RColorBrewer)  
   
  df %>%
    ggplot(aes(y=scr.rec_j, x=scr_j , colour=acq_j)) + 
    geom_point(alpha = .6) +  
     theme_farty() +
    scale_colour_gradientn(colours = brewer.pal(7, "Paired")) 
 
   
  df %>%
    ggplot(aes(y=scr.z.rec_j, x=scr_j , colour=sd_j)) + 
    geom_point(alpha = .6) +  
      theme_farty() +
    scale_colour_gradientn(colours = brewer.pal(7, "Paired")) 
   
  df %>%
    ggplot(aes(y=scr.z.rec_j, x=scr_j , colour=acq_j)) + 
    geom_point(alpha = .6) +  
       theme_farty() +
    scale_colour_gradientn(colours = brewer.pal(7, "Paired")) 
 
   df %>%
    ggplot(aes(y=sd_j, x=acq_j, colour=scr_j)) + 
    geom_point(alpha = .6) +  
     theme_farty() +
    scale_colour_gradientn(colours = brewer.pal(7, "Paired")) 
   
    df %>%
    ggplot(aes(y=sd_j, x=acq_j, colour=scr_j^2)) + 
    geom_point(alpha = .6) +  
     theme_farty() +
    scale_colour_gradientn(colours = brewer.pal(7, "Paired")) 
   
   
  df %>%
    ggplot(aes(y=scr.rec_j, x=sd_j, colour=acq_j)) + 
    geom_point(alpha = .6) +  
      theme_farty() +
    scale_colour_gradientn(colours = brewer.pal(7, "Paired")) 
  
  df %>%
    ggplot(aes(y=scr.z.rec_j, x=sd_j, colour=acq_j)) + 
    geom_point(alpha = .6) +  
    theme_farty() +
    scale_colour_gradientn(colours = brewer.pal(7, "Paired")) 
  
  df %>%
    ggplot(aes(y=scr.rec_j, x=acq_j, colour=sd_j)) + 
    geom_point(alpha = .6) +  
     theme_farty() +
    scale_colour_gradientn(colours = brewer.pal(7, "Paired")) 

  df %>%
    describe %>%
    kable(digits = 2)

```

```{r}


library(xlsx)

cbind(r_ptrn, df) %>% filter( acq_j == 1.5 |  acq_j == 3.5 |  acq_j == 4.5 ) %>% 
  arrange(desc(acq_j), sd_j, scr_j) %>% View()

cbind(r_ptrn, df) %>% filter( acq_j == 1.5 |  acq_j == 3.5 |  acq_j == 4.5 ) %>% 
  arrange(desc(acq_j), sd_j, scr_j) %>% write.xlsx(file= "acq4.xlsx")

cbind(r_ptrn, df) %>% 
  arrange(desc(acq_j), sd_j, scr_j) %>% write.xlsx(file= "acq4_full.xlsx")
```

