---
title: "R code for chapter 'Classical perspectives of controlling acquiescence with balanced scales'"
description: |
  This document presentes the code of the simulation described on the paper. .

author:
  - name: Ricardo Primi 
    url: http://lattes.cnpq.br/511415207991190
    affiliation: Universidade São Francisco, instituto Ayrton Senna
    affiliation_url: http://www.usf.edu.br/ppg/

  - name: Nelson Hauck-Filho  
    url:  
    affiliation: Universidade São Francisco
    affiliation_url:  http://www.usf.edu.br/ppg/
    

  - name: Felipe Valentini  
    url:  
    affiliation: Universidade São Francisco
    affiliation_url:  http://www.usf.edu.br/ppg/
        
  - name: Daniel D. Santos 
    url:  http://lattes.cnpq.br/1023570038147506
    affiliation: Universidade de São Paulo (USP-RP)
    affiliation_url: http://lepes.fearp.usp.br
 

bibliography: biblio.bib  
output: distill::distill_article


---

```{r include=FALSE}
library(distill)
library(radix)
```


#### Simulating balanced scale of 6 items (3 positively keyed and 3 negatively keyed)  

* Produces a dataset 9 items of 5 point likers scale.  
* Create all$5^9=1.953.125$ possible response patterns  


```{r echo=TRUE}
  5**9  

  library(crossdes)
  library(tidyverse)

  r_ptrn <- permutations(5, 9, repeats.allowed=TRUE)
  colnames(r_ptrn) <- paste("i", 1:9, sep="")
  
 # r_ptrn2 <-  permutations(5, 4, repeats.allowed=TRUE)

```
 
* Create mean and standard deviation within subjects and recode responses using two methods: (a) centering $c_{ij} = x_{ij}-acq_j$ and (b) $z_{ij} = (x_{ij}-acq_j)/sd_j$  


```{r echo=TRUE}
 
  acq_j <- apply(r_ptrn[, 1:6], 1, mean)
  sd_j <- apply(r_ptrn[, 1:6], 1, sd)
  
```

* Recode responses 
```{r}

 r_ptrn_rec <- apply(r_ptrn, 2, function(x){(x - acq_j)})
 r_ptrn_z_rec <- apply(r_ptrn_rec, 2, function(x){(x/sd_j)})

  hist(r_ptrn_rec[ , 4])
 
```



* Calculate scores. Note we reverse three last items as if they were negatively worded items.

```{r echo=TRUE}

  scr_j  <- r_ptrn %>% 
    as.data.frame %>%
    mutate_at(4:6,  funs(6-.)) %>%
    apply(1, mean, na.rm=TRUE)
  
  scr.rec_j  <- r_ptrn_rec %>% 
    as.data.frame %>%
    mutate_at(4:6,  funs(-1*.)) %>%
    apply(1, mean, na.rm=TRUE)
  
  scr.z.rec_j  <- r_ptrn_z_rec %>% 
    as.data.frame %>%
    mutate_at(4:6,  funs(-1*.)) %>%
    apply(1, mean, na.rm=TRUE)

```



* Create a tibble with scores acquiescence index  

```{r echo=TRUE}

 df <- tibble(
   scr_j,
   scr.rec_j,
   scr.z.rec_j,
   acq_j,
   sd_j
  )

```


#### Visualizing relationship among scores and acquiescence parameters

* How recoded scores (standardized and centered) are related to original scores
* How sd_j infuences transformation
* What is the relaioship between acq_j (mean) and standard deviation sd_j 




```{r, echo=TRUE}
  devtools::install_github('datarootsio/artyfarty')
  library(artyfarty)
  library(psych)
  library(knitr)
  library(ggplot2)
  library(RColorBrewer)  
   
  df %>%
    ggplot(aes(y=scr.rec_j, x=scr_j , colour=acq_j)) + 
    geom_point(alpha = .6) +  
     theme_farty() +
     scale_colour_gradientn(colours = brewer.pal(7, "Paired")) 
 
  
  df %>%
    ggplot(aes(y=scr.z.rec_j, x=scr_j , colour=sd_j)) + 
    geom_point(alpha = .6) +  
      theme_farty() +
    scale_colour_gradientn(colours = brewer.pal(7, "Paired")) 
  
   
  df %>%
    ggplot(aes(y=scr.z.rec_j, x=scr_j , colour=acq_j)) + 
    geom_point(alpha = .6) +  
       theme_farty() +
    scale_colour_gradientn(colours = brewer.pal(7, "Paired")) 
  
 
   df %>%
    ggplot(aes(y=sd_j, x=acq_j, colour=scr_j)) + 
    geom_point(alpha = .6) +  
     theme_farty() +
    scale_colour_gradientn(colours = brewer.pal(7, "Paired")) 
   
   
  df %>%
    ggplot(aes(y=scr.rec_j, x=sd_j, colour=acq_j)) + 
    geom_point(alpha = .6) +  
      theme_farty() +
    scale_colour_gradientn(colours = brewer.pal(7, "Paired")) 
  
  df %>%
    ggplot(aes(y=scr.z.rec_j, x=sd_j, colour=acq_j)) + 
    geom_point(alpha = .6) +  
    theme_farty() +
    scale_colour_gradientn(colours = brewer.pal(7, "Paired")) 
  
  df %>%
    ggplot(aes(y=scr.rec_j, x=acq_j, colour=sd_j)) + 
    geom_point(alpha = .6) +  
     theme_farty() +
    scale_colour_gradientn(colours = brewer.pal(7, "Paired")) 
  
   df %>%
    ggplot(aes(y=scr.rec_j , x=scr_j, colour=acq_j)) + 
    geom_point(alpha = .6) + 
    geom_abline(slope = 1, intercept = -3,  color = "black") +  
    theme_farty() +
    scale_colour_gradientn(colours = brewer.pal(7, "Paired")) 
   
   

  df %>%
    describe %>%
    kable(digits = 2)

```


```{r}
 source("http://www.labape.com.br/rprimi/R/cria_quartis.R")
  
  split_median <- function(x) {
    q <- quantile(x, probs =  .50, na.rm = TRUE )
    x <-  cut(x, q, ordered_result =FALSE, include.lowest = TRUE, 
      na.rm = TRUE,
    labels = c("<P50", ">P50")
    )
  }
   q <- quantile(x, probs =  .50, na.rm = TRUE )
  cut(df$acq_j, q, ordered_result =FALSE, include.lowest = TRUE, 
      na.rm = TRUE,
    labels = c("<P50", ">P50")

   split_median(df$acq_j)  
  
  
  df %>%
    ggplot(aes(y=scr.z.rec_j, x=scr_j , colour=acq_j)) + 
    geom_point(alpha = .6) +  
       theme_farty() +
    scale_colour_gradientn(colours = brewer.pal(7, "Paired")) + 
    facet_grid((acq_j <=1.5)~.)

```


```{r echo=TRUE}
 
 df <- as.data.frame(r_ptrn )

 
 df <- df %>% mutate(
    acq_j =  (i1 + i2 + i3 + i4 + i5 + i6) /6, 
    scr_j = ( i1 +i2 + i3 + (6-i4) + (6-i5) + (6-i6) )/6,
    scr_rec1_j = (
      (i1 - acq_j) + (i2 - acq_j) + (i3 - acq_j) +
      (6-(i4 - acq_j)) + (6-(i5 - acq_j)) + (6-(i6 - acq_j))
      ) /6,
      scr_rec2_j = ( ( ( (i1+ i2 + i3) / 3 ) - ( (i4+ i5 + i6) / 3 ) ) / 2   ) + 3
   )     
 
  df$sd_j <- sd_j

```


```{r}
  
  df %>%
    ggplot(aes(y=scr_rec1_j, x=scr_j , colour=acq_j)) + 
    geom_point(alpha = .6) +  
     theme_farty() +
    scale_colour_gradientn(colours = gray.colors(5, start = 0.98, end = 0.3, gamma = 2.2)) 
  

  df %>%
    ggplot(aes(y=scr_rec1_j, x=acq_j, colour=sd_j)) + 
    geom_point(alpha = .6) +  
     theme_farty() +
    scale_colour_gradientn(colours = gray.colors(5, start = 0.98, end = 0.3, gamma = 2.2)) 

  df %>%
    describe %>%
    kable(digits = 2)
```

```{r}
   
  df <- df %>% mutate(
  
    scr.z.rec_j = (
      (i1 - acq_j)/sd_j + (i2 - acq_j)/sd_j + (i3 - acq_j)/sd_j +
      -1*((i4 - acq_j)/sd_j ) + -1*((i5 - acq_j)/sd_j) + -1*((i6 - acq_j)/sd_j)
      ) /6
  
   )    
```


```{r}

  df %>%
    ggplot(aes(y= scr.z.rec_j  , x=scr_j, colour=acq_j, size = sd_j )) + 
    geom_point(alpha = .6) +  
     theme_farty() +
    scale_colour_gradientn(colours = gray.colors(5, start = 0.98, end = 0.3, gamma = 2.2)) 

 df %>%
    ggplot(aes(y= scr.z.rec_j  , x=scr_j, colour=acq_j)) + 
    geom_point(alpha = .6) +  
     theme_farty() +
    scale_colour_gradientn(colours = gray.colors(7, start = 0.98, end = 0.3, gamma = 2.2)) 

  
  
 df %>%
    ggplot(aes(y=sd_j , x=acq_j , colour=scr_j)) + 
    geom_point(alpha = .6) +  
     theme_farty() +
    scale_colour_gradientn(colours = gray.colors(5, start = 0.98, end = 0.3, gamma = 2.2)) 
  

  df %>%
    ggplot(aes(y= scr.z.rec_j, x=acq_j, colour=scr_j)) + 
    geom_point(alpha = .6) +  
     theme_farty() +
    scale_colour_gradientn(colours = gray.colors(5, start = 0.98, end = 0.3, gamma = 2.2)) 
  
   df %>%
    ggplot(aes(y= scr.z.rec_j, x=acq_j, colour=sd_j)) + 
    geom_point(alpha = .6) +  
     theme_farty() +
    scale_colour_gradientn(colours = gray.colors(5, start = 0.98, end = 0.3, gamma = 2.2)) 
   
    df %>%
    ggplot(aes(y= scr.z.rec_j, x=acq_j, colour=acq_j)) + 
    geom_point(alpha = .6) +  
     theme_farty() +
    scale_colour_gradientn(colours = gray.colors(5, start = 0.98, end = 0.3, gamma = 2.2)) 
  
    df %>%
    ggplot(aes(y= scr.z.rec_j, x=sd_j, colour=acq_j)) + 
    geom_point(alpha = .6) +  
     theme_farty() +
    scale_colour_gradientn(colours = gray.colors(5, start = 0.98, end = 0.3, gamma = 2.2)) 
 
    df %>%
    ggplot(aes(y= scr_rec1_j, x=sd_j, colour=acq_j)) + 
    geom_point(alpha = .6) +  
     theme_farty() +
    scale_colour_gradientn(colours = gray.colors(5, start = 0.98, end = 0.3, gamma = 2.2)) 
 
  
```


```{r eval=FALSE, echo=FALSE}

 df <- df %>%
   rowwise() %>%
   mutate(
     acq_j =  mean(c(i1, i2, i3, i4, i5, i6)),
     sd_j =  sd(c(i1, i2, i3, i4, i5, i6)),
     scr_j = mean(c(i1 +i2 + i3, (6-i4), (6-i5), (6-i6) )),
     scr_rec_j = ( (mean(c(i1, i2, i3) ) - mean( c(i4, i5, i6))) / 2)   + 3  
     )

```

